---

title: "Analysis"
author: "simplymathematics & jemceach"
date: "December 09, 2018"
output:
  html_document:
    highlight: monochrome
    theme: cosmo
  pdf_document: default
---

# Global Analysis of Internet Infrustucture & Equity

## ABSTRACT

As Data Scientists, we recognize that "big data" can be used to influence elections, spread hateful propaganda, and track every purchase and decision we make one. However, we believe that the Internet, as a whole provides, many positive economic outlets.  

We seek to quantify the positive effects of internet access on a global scale through our analysis of internet infrastructure and select social and economic indicators. 

We used data from the World Bank to measure this relationship and found XYZ....  

Our research process is outlined below for those interested in replicating our work.  

## R DEPENDENCIES

Replication requires the use of the following dependencies:

```{r dependencies, echo=TRUE, warning=FALSE, message=FALSE}
library(curl) 
library(XML) 
library(wbstats)
library(data.table)
library(tidyr)
library(dplyr)
library(knitr)
library(kableExtra)
suppressWarnings(source("indicators.R"))
```

## RESEARCH QUESTION

> Does internet access correlate with the chosen indicators of inequality?

    H0: Internet penetration rates are not correlated with the chosen equity indicators.
    HA: Internet penetrateion rates are correlated with the chosen equity indicator.

> Does the selected indicator have a stronger affect on internet access than the social equity indicators?
    
    H0 Economic indicators have a stronger effect than social measurements on Internet access.
    HA: Economic indicators do not have a stronger effect than social measurements on Internet access.


## INDICATORS 

We used the `curl` package to download our selected indicators from the [World Bank Indicator API Queries](https://datahelpdesk.worldbank.org/knowledgebase/articles/898599-indicator-api-queries). Documentation for this process can be found in the `indicator.R` file within this directory. 
 
The indicators, and their corresponding World Bank definitions, are outlined below:

1. Socio-Economic Indicators:
    - **SI.POV.GINI**: Gini index measures the extent to which the distribution of income...among individuals or households within an economy deviates from a perfectly equal distribution.
    - **NY.GDP.MKTP.KD.ZG:** Raw GDP. We chose raw GDP to see if internet infrastructure contributes more to social mobility than a generalized economy. 
    - **SE.ADT.LITR.ZS:** Literacy rate, adult total (% of people ages 15 and above)
    
     
2. Internet Indicators
    - **IT.NET.USER.ZS:** Internet users are individuals who have used the Internet (from any location) in the last 3 months. The Internet can be used via a computer, mobile phone, personal digital assistant, games machine, digital TV etc. This is the only dataset we had that was complete enough to be useful. While other indicators are available through the World Bank, they overselect for highly developed nations as they are able to measure and report such figures.

### Socio-Economic Indicators

#### GINI Index

**WB Note:** Gini index of 0 represents perfect equality, while an index of 100 implies perfect inequality."

NOTE: Look at this table from the [WB](http://wdi.worldbank.org/table/1.3). WB uses different reference year as data is not largely available during many periods. We may want to do something similiar. 


``` {r gini-table, echo=FALSE, warning=FALSE, message=FALSE}
kable(gini.wide, caption = 'GINI Values 1980 - 2016', format = "html")  %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = F, position = "c") %>%
  row_spec(row = 0:0, background = "lightgrey") %>%
  column_spec(column = 1, bold = T) 
```

#### Gross Domestic Product, Purchasing Power Parity (GDP, PPP) rates

``` {r gdp-table, echo=FALSE, warning=FALSE, message=FALSE}
kable(gdp.wide, caption = 'GDP, PPP (constant 2011 international $) ',format="html") %>%
kable_styling(bootstrap_options = c("striped", "hover"), full_width = F, position = "c") %>%
row_spec(row = 0:0, background = "lightgrey") %>%
column_spec(column = 1, bold = T) 
```

### TODO Literacy

### Internet Indicators
```{r}
internet <- read.csv("internet/internet_over_time.csv")
plot(internet$annual.mean~internet$years)
model <- lm(internet$annual.mean~internet$years)
abline(model)
summary(model)
```
The average speed of internet access around the world grows by a factor of 2.652 every year! Furthermore, this model has an $R^2$ value of .998, meaning that global internet access speeds are almost entirely a function of time.


## ANALYSIS 

First, I'd compare gini vs internet
```{r}
plot(gini.annual~internet$annual.mean)
model <- lm(gini.annual~internet$annual.mean)
model2 <- lm(gini.annual~gdp.annual)
summary(model)
summary(model2)
```

Repeat for lit vs internet (make sure to use correct indicator i.e. literacy$)
```{r}
plot(lit.annual~internet$annual.mean)
model <- lm(lit.annual~internet$annual.mean)
model2 <- lm(lit.annual~gdp.annual)
summary(model)
summary(model2)

```

```{r}
plot(internet$annual.mean ~ gdp.annual)
model <- lm(internet$annual.mean ~gdp.annual)
abline(model)
summary(model)
```

# CONCLUSION

GDP is a stronger indicator of literacy or wealth inequality than internet access rates are. 

Challenge: Finding an internet indicator that had enough data points. Finding a social indicator that was a good comparison across countries. Finding a measure of social equity was hard for similar reasons.

